---
title: "Tricho_Colim_DESeq2"
author: "Nina Yang, PhD Candidate, Hutchins Lab, USC"
adapted by: "Kyla Kelly, PhD Candidate, Hutchins Lab, USC"
date: "9/28/2021"
updated: "3/10/2022"
output: html_document
---

## Background

This document outlines the steps taken to analyze the RNASeq data using DESeq2 

## Load libraries

**1. Install (if required) & upload packages into your current session**

```{r install-packages, eval = TRUE, message = FALSE, warning = FALSE}
if(!require(dplyr)){install.packages("dplyr")} 
if(!require(tidyverse)){install.packages("tidyverse")} 
if(!require(ggpubr)){install.packages("plyr")}
if(!requireNamespace("BiocManager", quietly = TRUE))
  install.packages("BiocManager")
BiocManager::install("DESeq2")
```

**2. Upload the package into your current session**
```{r library, eval = TRUE, message = FALSE, warning = FALSE}
library(DESeq2)
library(tidyverse) 
library(dplyr) 
library(plyr)
library(ggplot2)
```

## RNASeq Analysis with DESeq2

Remove duplicate KOs

```{r ddply, eval = TRUE, message = FALSE, warning = FALSE}
# Gene counts were assigned to Kegg IDs and the duplicate KOs in the dataset will be merged.
# setwd("/Users/kylakelly/Documents/USC/research/seq/tricho/deseq2/")

# did some work in excel to rename columns, and find and replace blanks KOs with "NA"
df <- read.csv("deseq2/Tricho_fewarm_counts_stranded_v2.csv") 

df$KO <- ifelse(is.na(df$KO), df$gene_id, df$KO)
df = df[,2:20]
names(df)[which(names(df) == "KO")] <- "gene_id"

colnames(df)
summary(df) 
head(df)
dim(df) # there are 15,971 entries
ddDF <- ddply(df,"gene_id",numcolwise(sum))
dim(ddDF) # now there are  15,125 entries
write.csv(ddDF,"deseq2/tricho_counts_deduped_forDESeq2.csv", row.names = FALSE)
```

Normalize data for subsequent analysis. After removing duplicates, the counts file will be trimmed to remove rows where the average of each treatment for a given gene is less than 5 (we keep genes where the average of at least one treatment is greater than 5). 

```{r trim, eval = TRUE, message = FALSE, warning = FALSE}
# trimmed counts is the data where genes with low counts have been removed

trim_df <- read_csv("deseq2/tricho_counts_deduped_forDESeq2.csv")
head(trim_df)
colnames(trim_df)
dim(trim_df)  # 15,125 x 19

# calculate rowmeans
mean.L22 <- as.numeric(rowMeans(trim_df[, 2:4]))
mean.R22 <- as.numeric(rowMeans(trim_df[, 5:7]))
mean.L27 <- as.numeric(rowMeans(trim_df[, 8:10]))
mean.R27 <- as.numeric(rowMeans(trim_df[, 11:13]))
mean.L32 <- as.numeric(rowMeans(trim_df[, 14:16]))
mean.R32 <- as.numeric(rowMeans(trim_df[, 17:19]))

# create a dataframe only with rowmeans based on treatment
trim_df1 <- as.data.frame(cbind(mean.L22, mean.R22, mean.L27, mean.R27, mean.L32, mean.R32))
colnames(trim_df1)
summary(trim_df1)

trim_df2 <- cbind(trim_df$gene_id, trim_df1)
summary(trim_df2)
colnames(trim_df2)
names(trim_df2)[1] <- "gene_id"
head(trim_df2)
summary(trim_df2)
dim(trim_df2) # 15,125 x 7

# temp effort to check against manual effort
# write.csv(trim_df2,"deseq2/tricho_trimming_temp.csv", row.names = FALSE)

# filter this rowMeans dataframe so that all rows where even one column is less than 5 is removed
# I think what's actually happening is that if ALL rows are <5, they're removed. Idk which one I really want 
trim_rows <- trim_df2 %>% dplyr::select(gene_id, starts_with("mean")) %>% filter_if(is.numeric, any_vars(. >= 5)) # should have 2843. Checked against a manual effort.

summary(trim_rows)
dim(trim_rows) # 4475 x 7

trimmed <- left_join(trim_rows, trim_df, by="gene_id")
head(trimmed)

final_trim <- trimmed %>% dplyr::select(gene_id, !(starts_with("mean")))
head(final_trim)
summary(final_trim)

write.csv(final_trim, file = "deseq2/tricho_trimmed_forDESeq2.csv", row.names = TRUE)

```

Setting up the matrix for DESeq2 and generating normalized datasets.


```{r normalize, eval = TRUE, message = FALSE, warning = FALSE}

cts <- read_csv("deseq2/tricho_trimmed_forDESeq2.csv")
colnames(cts);
cts <- cts[,3:20]
colnames(cts)

# column data is the metadata for the dataset
## the data has Sample, Treatment, Diel, and Design (combination of Treatment and Diel)
col <- read.csv("deseq2/tricho_coldata_DESeq2.csv", header = TRUE)

# set up DESeq2 matrix for Design
dds <- DESeqDataSetFromMatrix(countData = cts, colData = col, design = ~ Design)

# get normalized counts: not for DESeq but important for downstream visualization (DESeq2 input is non-normalized counts data because it undergoes normalization as a part of the pipeline). We will be using normalized counts for visualizations of biomarker genes.
# https://hbctraining.github.io/DGE_workshop/lessons/02_DGE_count_normalization.html
# note that DESeq2 recommend vst transformed!!
ddsNorm <- estimateSizeFactors(dds)
sizeFactors(ddsNorm)
normalized_counts <- counts(ddsNorm, normalized=TRUE)
write.csv(as.data.frame(normalized_counts), file="deseq2/tricho_normcounts_forDGE.csv")

# plot a PCA to check the data 
vstcounts <- vst(dds, blind=FALSE)
write.csv((assay(vstcounts)), file="deseq2/tricho_vstcounts_forDGE.csv")
head(assay(vstcounts), 3)
plotPCA(vstcounts, intgroup=c("Treatment", "Temp"))
ggsave("deseq2/PCA_vstcounts.pdf", height=75, width=120, units='mm', dpi=500)
write.csv((assay(vstcounts)), file="deseq2/tricho_vstcounts_forDGE.csv")
# I added the line of code above but idk if I actually need it 
rld <- rlog(dds, blind=FALSE)
plotPCA(rld, intgroup=c("Treatment", "Design"))
ggsave("deseq2/PCA_rdl.pdf", height=75, width=120, units='mm', dpi=500)
write.csv((assay(rld)), file="deseq2/tricho_rldcounts_forDGE.csv")

# don't use normalized counts (norm counts are used for comparison and other calculations)

```

# Pairwise comparisons for deseq2
```{r}

ddsObj <- DESeq(dds)
res <- results(ddsObj, alpha = 0.05) # filtering set to alpha 0.05
sum(res$padj < 0.05, na.rm=TRUE) # this pulls the number of genes with padj less than 0.05 (total) 
# 1246

# Replete v Limited by temperature 
## Replete v Limited Replete means that positive is upregulated for Replete relative to Limited

#R22 vs L22
df <- results(ddsObj, alpha = 0.05, contrast = c("Design", "R22", "L22"))
summary(df)
# 341 up
# 327 down
# 521 low counts 
# no outliers in any of these treatments here and below

# write the file as a csv
write.csv(as.data.frame(df),
          file="deseq2/outputs/R22vL22_DESeqresults.csv")


#R27 vs L27
df <- results(ddsObj, alpha = 0.05, contrast = c("Design", "R27", "L27"))
summary(df)
# 917 up
# 759 down
# 174 low counts 

# write the file as a csv
write.csv(as.data.frame(df),
          file="deseq2/outputs/R27vL27_DESeqresults.csv")

#R32 vs L32
df <- results(ddsObj, alpha = 0.05, contrast = c("Design", "R32", "L32"))
summary(df)
# 414 up
# 314 down
# 0 low counts 

# write the file as a csv
write.csv(as.data.frame(df),
          file="deseq2/outputs/R32vL32_DESeqresults.csv")


## comparing sub (22) and supra (32) relative to optimum (27)

#L22 vs L27
## extracts only limited 22 vs limited 32
## L22 vs L27 means that positive is upregulated for L22 relative to L27
df <- results(ddsObj, alpha = 0.05, contrast = c("Design", "L22", "L27"))
summary(df)
# 672 up
# 552 down
# 694 low counts 

# write the file as a csv
write.csv(as.data.frame(df),
          file="deseq2/outputs/L22vL27_DESeqresults.csv")

#L27 vs L32
## L27 v L32 means that positive is upregulated for L27 relative to L32
df <- results(ddsObj, alpha = 0.05, contrast = c("Design", "L27", "L32"))
summary(df)
# 613 up
# 652 down
# 521 low counts 

# write the file as a csv
write.csv(as.data.frame(df),
          file="deseq2/outputs/L27vL32_DESeqresults.csv")

#R22 vs R27
## R22 v R27 means that positive is upregulated for R22 relative to R27
df <- results(ddsObj, alpha = 0.05, contrast = c("Design", "R22", "R27"))
summary(df)
# 438 up
# 485 down
# 434 low counts 

# write the file as a csv
write.csv(as.data.frame(df),
          file="deseq2/outputs/R22vR27_DESeqresults.csv")

#R27 vs R32
## R27 v R32 means that positive is upregulated for R27 relative to R32
df <- results(ddsObj, alpha = 0.05, contrast = c("Design", "R27", "R32"))
summary(df)
# 316 up
# 223 down
# 868 low counts 

# write the file as a csv
write.csv(as.data.frame(df),
          file="deseq2/outputs/R27vR32_DESeqresults.csv")


# sub (27) vs supra (32)

#R22 vs R32
## R22 v R32 means that positive is upregulated for R22 relative to R32
df <- results(ddsObj, alpha = 0.05, contrast = c("Design", "R22", "R32"))
summary(df)
# 610 up
# 565 down
# 0 low counts 

# write the file as a csv
write.csv(as.data.frame(df),
          file="deseq2/outputs/R22vR32_DESeqresults.csv")

#L22 vs L32
## L22 v L32 means that positive is upregulated for L22 relative to L32
df <- results(ddsObj, alpha = 0.05, contrast = c("Design", "L22", "L32"))
summary(df)
# 683 up
# 650 down
# 347 low counts 

# write the file as a csv
write.csv(as.data.frame(df),
          file="deseq2/outputs/L22vL32_DESeqresults.csv")

```