---
title: "Tricho.GO.annot.compile"
output: html_document
date: "2023-01-04"
https://github.com/yang-nina/Crocosphaera-Fe-P-Colimitation/blob/main/Croco_Annotation/Croco_Colim_annotation_compile.Rmd 
---

```{r}
library(BiocManager)
library(tidyverse)
library(GO.db)
library(dplyr)
library(AnnotationDbi)
```

## GO Annotation

This script will take the results from the DIAMOND blastx run, subsequent BLAST2GO mapping with InterProScan annotation, and Uniprot annotation to functionally annotate the Crocosphaera WH0005 genome with Gene Ontology (GO) terms.

**Blast**
This script will take the results from the DIAMOND blastx run, subsequent BLAST2GO mapping with InterProScan annotation, and Uniprot annotation to functionally annotate the Crocosphaera WH0005 genome with Gene Ontology (GO) terms.

Blastx using DIAMOND took place on Feb 3rd, 2023 using the NCBI nr database downloaded 1/2/23 at 7:36pm. 

```{r blastx, eval = TRUE, message = FALSE, warning = FALSE}
blast <- read_tsv(file = "Tery_IMS101_diamond_annot_new.tsv") # this is the output from DIAMOND blastx
head(blast)
colnames(blast) <- c("seqName", "tophit", "pident", "length", "mismatch", "gapopen", "qstart", "qend", "sstart", "send", "evalue", "bitscore",  "qlen", "slen") # rename columns
summary(blast)
head(blast)
dim(blast) # 5073 x 16
```

**Uniprot**
Uniprot mapping took place on March 1st, 2023 

Refseq Protein IDs were Mapped to UniProtKB IDs


```{r uniprot_blast, eval = TRUE, message = FALSE, warning = FALSE}
uniprot_1 <- read_tsv("RefSeq_Protein-to-UniProtKB.tsv", col_names = TRUE)
uniprot_1 <- uniprot_1[,c(1,3:10)] # select relevant columns
colnames(uniprot_1) <- c("tophit", "uniprotkb_entry", "status", "protein_names", "gene_names", "organism", "length", "gene_ontology", "go_ids")
head(uniprot_1)
dim(uniprot_1) # 20
```

EMBL / GENBANK / DDBJ / CDS / IDs were mapped to UniProtKB IDs

```{r uniprot_refseq, eval = TRUE, message = FALSE, warning = FALSE}
uniprot_2 <- read_tsv("EMBL-GenBank-DDBJ_CDS-to-UniProtKB.tsv", col_names = TRUE)
uniprot_2 <- uniprot_2[,c(1,3:10)]
colnames(uniprot_2) <- c("tophit", "uniprotkb_entry", "status", "protein_names", "gene_names", "organism", "length", "gene_ontology", "go_ids")
head(uniprot_2)
dim(uniprot_2) #179
```

UniProtKB_AC-ID-to-UniProtKB-Swiss-Prot

```{r uniprot_refseq, eval = TRUE, message = FALSE, warning = FALSE}
uniprot_3 <- read_tsv("UniProtKB_AC-ID-to-UniProtKB-Swiss-Prot.tsv", col_names = TRUE)
uniprot_3 <- uniprot_3[,c(1,3:10)]
colnames(uniprot_3) <- c("tophit", "uniprotkb_entry", "status", "protein_names", "gene_names", "organism", "length", "gene_ontology", "go_ids")
head(uniprot_3)
dim(uniprot_3) # 335
```
    
UniProtKB_AC-ID-to-UniProtKB

```{r uniprot_refseq, eval = TRUE, message = FALSE, warning = FALSE}
uniprot_4 <- read_tsv("UniProtKB_AC-ID-to-UniProtKB.tsv", col_names = TRUE)
uniprot_4 <- uniprot_4[,c(1,3:10)]
colnames(uniprot_4) <- c("tophit", "uniprotkb_entry", "status", "protein_names", "gene_names", "organism", "length", "gene_ontology", "go_ids")
head(uniprot_4)
dim(uniprot_4) # 335
```

**Compile Uniprot Results**

```{r compile_uniprot, eval = TRUE, message = FALSE, warning = FALSE}
Uniprot_results <- bind_rows(uniprot_1, uniprot_2, uniprot_3, uniprot_4) # combines uniprot results
Uniprot_results <- unique(Uniprot_results) # I think mine were already all unique
Uniprot_results$go_ids <- gsub(" ", "", Uniprot_results$go_ids) # remove unwanted spaces
head(Uniprot_results)
dim(Uniprot_results) #534 x 9
Uniprot_results$gene_ontology <- gsub(" \\[.*?\\]", "", Uniprot_results$gene_ontology) #remove unwanted information within brackets to streamline compilation later on
Uniprot_results$gene_ontology <- gsub("; ", ";", Uniprot_results$gene_ontology) #remove unwanted characters
head(Uniprot_results)
nrow(filter(Uniprot_results, grepl("GO:",go_ids))) #Genes with GO terms #439
dim(Uniprot_results) # 534

```

**Blast2GO** 
Nina created this one for me this time

From Nina: Of 5,924 coding regions, 5,065 returned accessions via DIAMOND blastx. BLAST2GO GO mapping and IPS annotation identified 2,866 accessions with GO annotation.
# this is from Nina
# some of my "descriptions" are missing names

```{r blast2go_merge, eval = TRUE, message = FALSE, warning = FALSE}
B2G_results <- read_csv("blast2go_table_TeryISM101_new.csv", col_names = TRUE)
head(B2G_results)
B2G_results <- B2G_results[,c(3:5, 7:8,10:11, 15:16)] # select the relevant columns
head(B2G_results)
colnames(B2G_results) <- c("seqName",  "tophit", "length",  "eValue", "simMean", "GO_IDs", "GO_names", "InterPro_GO_IDs", "InterPro_GO_names") # rename the relevant columns

B2G_results$GO_IDs <- gsub(" ", "", B2G_results$GO_IDs) # remove unwanted spaces
B2G_results$InterPro_GO_IDs <- gsub(" ", "", B2G_results$InterPro_GO_IDs) # remove unwanted spaces
head(B2G_results)

B2G_results$InterPro_GO_IDs <- gsub("noGOterms", "NA", B2G_results$InterPro_GO_IDs) # replace with NA
B2G_results$InterPro_GO_names <- gsub("no GO terms", "NA", B2G_results$InterPro_GO_names) # replace with NA
B2G_results$InterPro_GO_IDs <- gsub("noIPSmatch", "NA", B2G_results$InterPro_GO_IDs) # replace with NA
B2G_results$InterPro_GO_names <- gsub("no IPS match", "NA", B2G_results$InterPro_GO_names) # replace with NA
head(B2G_results)
dim(B2G_results) #5284
nrow(filter(B2G_results, grepl("GO:",GO_IDs))) # 2222 GO
nrow(filter(B2G_results, grepl("GO:",InterPro_GO_IDs))) # 5726 IPS

B2G_results_clean = B2G_results %>% 
  unite(concat_IDs, c(GO_IDs, InterPro_GO_IDs), sep = ";", remove = FALSE) %>%
  unite(concat_ontology, c(GO_names, InterPro_GO_names), sep = ";", remove = FALSE)  # merge GO and IPS IDs
head(B2G_results_clean)
B2G_results_clean$concat_ontology <- gsub("; ", ";", B2G_results_clean$concat_ontology) # remove unwanted space after ;
head(B2G_results_clean$concat_ontology)
B2G_results_clean$concat_IDs <- sapply(B2G_results_clean$concat_IDs, function(x) paste(unique(unlist(str_split(x,";"))), collapse = ";")) # remove duplicated IDs after merge
B2G_results_clean$concat_ontology <- sapply(B2G_results_clean$concat_ontology, function(x) paste(unique(unlist(str_split(x,";"))), collapse = ";")) # remove duplicated terms after merge
head(B2G_results_clean)

B2G_results_clean$concat_IDs <- gsub("F:", "", B2G_results_clean$concat_IDs) # remove unwanted characters for downstream processing
B2G_results_clean$concat_ontology <- gsub("F:", "", B2G_results_clean$concat_ontology) # remove unwanted characters for downstream processing
B2G_results_clean$concat_IDs <- gsub("P:", "", B2G_results_clean$concat_IDs) # remove unwanted characters for downstream processing
B2G_results_clean$concat_ontology <- gsub("P:", "", B2G_results_clean$concat_ontology) # remove unwanted characters for downstream processing
B2G_results_clean$concat_IDs <- gsub("C:", "", B2G_results_clean$concat_IDs) # remove unwanted characters for downstream processing
B2G_results_clean$concat_ontology <- gsub("C:", "", B2G_results_clean$concat_ontology) # remove unwanted characters for downstream processing
B2G_results_clean$concat_IDs <- gsub("NA;", "", B2G_results_clean$concat_IDs) # remove unwanted characters for downstream processing
B2G_results_clean$concat_ontology <- gsub("NA;", "", B2G_results_clean$concat_ontology) # remove unwanted characters for downstream processing

dim(B2G_results_clean) # 5284
nrow(filter(B2G_results_clean, grepl("GO:",concat_IDs))) #2622 accessions total (same as GO mapping) 
head(B2G_results_clean)
tail(B2G_results_clean)
dim(B2G_results_clean) # 5284   11
```

**Find unique and overlapping GO terms**
Generate lists of GO terms for each method

```{r unique_overlap, eval = TRUE, message = FALSE, warning = FALSE}

Uniprot_GO <- dplyr::select(Uniprot_results, tophit, go_ids)
splitted <- strsplit(as.character(Uniprot_GO$go_ids), ";") # split into multiple GO ids
gene_ontology <- data.frame(v1 = rep.int(Uniprot_GO$tophit, sapply(splitted, length)), v2 = unlist(splitted)) # list all genes with each of their GO terms in a single row
head(gene_ontology)
colnames(gene_ontology) <- c("gene_id", "GO.ID")
head(gene_ontology)
Uniprot.GOterms <- dplyr::select(gene_ontology, gene_id, GO.ID)
Uniprot.GOterms$GO.ID<- as.character(Uniprot.GOterms$GO.ID)
Uniprot.GOterms[Uniprot.GOterms == 0] <- "unknown"
Uniprot.GOterms$GO.ID <- replace_na(Uniprot.GOterms$GO.ID, "unknown") # replace NAs
head(Uniprot.GOterms)
Uniprot.GOterms$GO.ID <- as.factor(Uniprot.GOterms$GO.ID)
Uniprot.GOterms$gene_id <- as.factor(Uniprot.GOterms$gene_id)
Uniprot.GOterms$GO.ID <- gsub(" ", "", Uniprot.GOterms$GO.ID) # remove unwanted spaced
Uniprot.GOterms <- unique(Uniprot.GOterms)
nrow(Uniprot.GOterms) # 1748 rows / terms
head(Uniprot.GOterms)

B2G_GO <- dplyr::select(B2G_results_clean, tophit, concat_IDs)
splitted <- strsplit(as.character(B2G_GO$concat_IDs), ";") # split into multiple GO ids
gene_ontology <- data.frame(v1 = rep.int(B2G_GO$tophit, sapply(splitted, length)), v2 = unlist(splitted)) # list all genes with each of their GO terms in a single row
dim(gene_ontology) # 35519
colnames(gene_ontology) <- c("gene_id", "GO.ID")
gene_ontology <- filter(gene_ontology, !is.na(gene_id)) # remove entries where gene_id (accession) is NA
dim(gene_ontology) #35309 
B2G.GOterms <- dplyr::select(gene_ontology, gene_id, GO.ID)
B2G.GOterms$gene_id <- as.character(B2G.GOterms$gene_id)
B2G.GOterms$GO.ID <- as.character(B2G.GOterms$GO.ID)
B2G.GOterms[B2G.GOterms == 0] <- "unknown"
B2G.GOterms$GO.ID <- tidyr::replace_na(B2G.GOterms$GO.ID, "unknown")
B2G.GOterms$GO.ID <- gsub("NA", "unknown", B2G.GOterms$GO.ID)
head(B2G.GOterms)
B2G.GOterms$GO.ID <- as.factor(B2G.GOterms$GO.ID)
B2G.GOterms$gene_id <- as.factor(B2G.GOterms$gene_id)
B2G.GOterms$GO.ID <- gsub(" ", "", B2G.GOterms$GO.ID) # remove unwanted spaces
B2G.GOterms <- unique(B2G.GOterms)
nrow(B2G.GOterms) # 34565 terms
head(B2G.GOterms) 

# Find intersections and unique results for each methods
UB <- dplyr::intersect(B2G.GOterms, Uniprot.GOterms) # Blast2GO and Uniprot intersection
nrow(UB) # 1719 terms intersect
Uunique <- dplyr::setdiff(Uniprot.GOterms, B2G.GOterms) # Uniprot unique
nrow(Uunique) # 29 terms (this makes sense, I didn't have THAT many uniprot genes to begin with)
Bunique <- dplyr::setdiff(B2G.GOterms, Uniprot.GOterms) # Blast unique
nrow(Bunique) # 32846 terms
```

**Merge Annotations**

Match tophits with description

```{r merge_annot, eval = TRUE, message = FALSE, warning = FALSE}
tricho_annot <- left_join(blast, B2G_results_clean, by="seqName")
head(tricho_annot)
tail(tricho_annot)
dim(tricho_annot) #5073 x 22

tricho_annot <- dplyr::select(tricho_annot, seqName, tophit.x, length.x, evalue, bitscore, simMean, concat_IDs, concat_ontology) 
#tricho_annot <- tricho_annot[,c(1:2,4,11,12,17:18,20)] # select the relevant columns

tricho_annot <- dplyr::rename(tricho_annot, "tophit" = "tophit.x")
head(tricho_annot)
tricho_annot <- left_join(tricho_annot, Uniprot_results, by ="tophit")
head(tricho_annot)
dim(tricho_annot) # increase initially in annotation rows due to adding uniprot results
# 5087 x 16

tricho_annot$GO <- paste(tricho_annot$concat_IDs, tricho_annot$go_ids, sep=';') # generate new column with concatenated GO IDs
head(tricho_annot)
tricho_annot$GO_terms <- paste(tricho_annot$concat_ontology, tricho_annot$gene_ontology, sep=';') # generate new column with concatenated GO IDs
# the following command with -c means these columns are excluded
tricho_annot <- dplyr::select(tricho_annot,-c("concat_IDs", "concat_ontology", "gene_ontology", "go_ids", "length", "uniprotkb_entry", "status", "protein_names", "gene_names", "organism"))
# tricho_annot <- tricho_annot[,c(1:6,17:18)] # select the relevant columns

head(tricho_annot)
tail(tricho_annot)
colnames(tricho_annot) <- c("gene_id", "description", "length","eValue", "bitscore","simMean", "GO_IDs","GO_terms")
names(tricho_annot)
head(tricho_annot)
tail(tricho_annot)
dim(tricho_annot) # 5087     8

# need to remove duplicate GO terms after merging Uniprot and blast 
tricho_annot$GO_IDs <- sapply(tricho_annot$GO_IDs, function(x) paste(unique(unlist(str_split(x,";"))), collapse = ";"))
tricho_annot$GO_terms <- sapply(tricho_annot$GO_terms, function(x) paste(unique(unlist(str_split(x,";"))), collapse = ";"))
tricho_annot$GO_IDs <- gsub(";NA", "", tricho_annot$GO_IDs)
tricho_annot$GO_terms <- gsub(";NA", "", tricho_annot$GO_terms)
tricho_annot$GO_IDs <- gsub("NA;", "", tricho_annot$GO_IDs)
tricho_annot$GO_terms <- gsub("NA;", "", tricho_annot$GO_terms)
dim(tricho_annot) # 5087     8
head(tricho_annot)
median(tricho_annot$eValue) # 1.75e-152
median(tricho_annot$bitscore) #432
df <- tricho_annot %>%
  filter(GO_IDs != "NA") # remove entries without GO terms
head(df)
dim(df) # 2637    8

```

```{r}

kegg <- read_csv("TeryIMS101_kofamscan.csv", col_names = TRUE) # read in KEGG annotation (this is based on ...)
head(kegg)
dim(kegg) # 1971

#rename this to annotate with KO
colnames(kegg)[1] <- "gene_id"
head(kegg)

# remove the * at the beginning of every id
kegg$gene_id <- sub("^\\*", "", kegg$gene_id)
head(kegg)
dim(kegg) # 1971
names(kegg)[6] = "gene_name"
KO = kegg$KO
kegg = cbind(kegg, KO)
names(kegg)[1] = "id"
names(kegg)[2] = "gene_id"

kegg = kegg[,c(1:2,6:7)]

# doing this silly thing so the data types are the same 
write.csv(kegg, file="kegg_forR.csv")
write.csv(df, file="df_forR.csv")

kegg = read.csv(file = "kegg_forR.csv")
kegg = kegg[,-1]
head(kegg)
dim(kegg) # 1971
df = read.csv(file="df_forR.csv")
df = df[,-1]
names(df)[1] = "id"
head(df)
dim(df) #2637

# KEGG annotation
KO_annot <- left_join(df, kegg, by="id")
head(KO_annot)
dim(KO_annot) # 2764, only the genes that have GO annotation (but this is more than the previous so im confused)
#maybe there are unique KOs and unique GOs --> yes
KO_annot = KO_annot[,c(1:9,13:14)]

write.csv(KO_annot, "tricho_annot_compiled_final.csv", row.names = TRUE)
```


**Build GO ID / TERM Annotation Database for Enrichment Analysis**

This takes the score with a list of GO IDs separated by ";" and creates an individual row for each GO ID and the associated gene_id. This will be used for GO enrichment analysis after DESeq2.

```{r split_annot, eval = TRUE, message = FALSE, warning = FALSE}

# This is used for the "universe" for enrichment analysis. Split the collated GO IDs into individual rows associated by gene_id. The gene_id list will include merged KOs (combine replicate KOs to generate "unique"" KOs) and numeric (assigned) gene_ids.

df <- read_csv("tricho_annot_compiled_final.csv", col_names = TRUE) # load the annotation file
head(df)
tail(df)
colnames(df)

df$gene_id <- ifelse(is.na(df$gene_id), df$description, df$gene_id)

df <- (df[, c(2, 8:12)]) # extract relevant columns
dim(df) # 2764
head(df)

# instead of NA in the KOs, it needs to be none
df$KO = ifelse(is.na(df$KO),"none", df$KO)

# split the dataset by KOs that are "none" and everything else
none_df = dplyr::filter(df, KO == "none")
head(none_df)
tail(none_df)
dim(none_df) #1039

ids <- none_df$gene_id # replace "none" with the id
head(ids)
none_df <- cbind(none_df, ids)
head(none_df)
tail(none_df)
colnames(none_df)
none_df <- dplyr::select(none_df, 1:4)
head(none_df)
names(none_df)[4] <- "KO" # rename as "KO" # this column includes the accession codes from blastx and not actual KO identifiers
colnames(none_df)

# for those that are KO-annotated:
KO_df = dplyr::filter(df, KO != "none")
head(KO_df)
dim(KO_df) # 1725
               
IDs_KO_df <- KO_df %>% group_by(KO) %>% 
  mutate(GO_IDs = paste(GO_IDs, collapse = ";")) %>% distinct(KO, .keep_all = TRUE) # this gets unique KOs and merges GO_IDs

dim(IDs_KO_df) # 1313 x 6
head(IDs_KO_df)

terms_KO_df <- KO_df %>% group_by(KO) %>% 
  mutate(GO_terms = paste(GO_terms, collapse = ";")) %>% distinct(KO, .keep_all = TRUE) # this gets unique KOs and merges terms

dim(terms_KO_df) # 1313 x 6
head(terms_KO_df)

IDs_terms <- dplyr::setdiff(IDs_KO_df$KO, terms_KO_df$KO)
IDs_terms <- dplyr::intersect(IDs_KO_df$KO, terms_KO_df$KO)
nrow(IDs_terms) # this should be null (none because the list should be identical)
# NULL

# combine these two such that the GO_IDs and GO_terms are replaced with the correct merged lists.
head(IDs_KO_df)
head(terms_KO_df)

IDs <- dplyr::select(IDs_KO_df, 1, 2, 4, 6)
head(IDs)
dim(IDs) #1323

terms <- dplyr::select(terms_KO_df, 1, 3)
head(terms)
dim(terms) #1323

kegg_annot <- left_join(IDs, terms, by = "KO") 
head(kegg_annot)
tail(kegg_annot)

# LEFT OFF HERE 
kegg_annot <- dplyr::select(kegg_annot, -c(5)) # remove duplicate id column
names(kegg_annot)[1] <- "gene_id"
names(kegg_annot)[3] <- "KO_id"
names(kegg_annot)[4] <- "KO_acc"
head(kegg_annot)
tail(kegg_annot)
kegg_annot <- kegg_annot[, c(1, 3, 4, 2, 5)] # reorder columns
head(kegg_annot)
dim(kegg_annot) #1313

# now we have to combine the merged KOs and GO terms with the genes that do not have KO annotation
colnames(none_df)[1] <- "gene_id" # rename column 1 to gene_id
colnames(none_df)[4] <- "KO_acc" # rename column 1 to gene_id

KO_id = none_df$gene_id
none_df <- cbind(none_df, KO_id)
head(none_df)

none_df <- none_df[, c(1, 5, 4, 2, 3)] # reorder columns
head(none_df)
head(kegg_annot)

#have to do this weird read/write business because I kept getting an error saying something about the types of data in the columns not being the same 
write_csv(none_df, "none_df.csv")
write_csv(kegg_annot, "kegg_annot.csv")

none_df = read_csv("none_df.csv")
kegg_annot = read_csv("kegg_annot.csv")
head(none_df)
head(kegg_annot)

final_KO_annot <- rbind(kegg_annot, none_df)
head(final_KO_annot)
tail(final_KO_annot)

dim(final_KO_annot) # final annotation worksheet for Crocosphaera files where duplicated KOs are merged; also the "nones" are replaced with blastx accession ids (2352 x 5 dim)

# need to remove duplicate GO terms after merging rows based on KOs
final_KO_annot$GO_IDs <- sapply(final_KO_annot$GO_IDs, function(x) paste(unique(unlist(str_split(x,";"))), collapse = ";"))
final_KO_annot$GO_terms <- sapply(final_KO_annot$GO_terms, function(x) paste(unique(unlist(str_split(x,";"))), collapse = ";"))
final_KO_annot = final_KO_annot[,1:5]
head(final_KO_annot)
tail(final_KO_annot)
dim(final_KO_annot) #2352, same number as above (deduping within the GO terms, not by rows, so this number should be the same as above)

splitted <- strsplit(as.character(final_KO_annot$GO_IDs), ";") # split into multiple GO ids
head(splitted)

df_split <- data.frame(v1 = rep.int(final_KO_annot$KO_id, sapply(splitted, length)), v2 = unlist(splitted)) # list all genes with each of their GO terms in a single row
df_split2 <- data.frame(v1 = rep.int(final_KO_annot$KO_acc, sapply(splitted, length)), v2 = unlist(splitted)) # list all genes with each of their GO terms in a single row
head(df_split, 10)
tail(df_split, 10)
dim(df_split) # 30691
head(df_split2, 10)
tail(df_split2)
dim(df_split2) #30691

colnames(df_split) <- c("gene_id", "GO.ID") # rename columns
colnames(df_split2) <- c("KO", "GO.ID")
head(df_split) # check colnames
head(df_split2)
bind <- cbind(df_split, df_split2)
head(bind)
dim(bind) #30691
bind_df <- bind[1:3]
head(bind_df)
tail(bind_df)

# FROM NY:
# I need to annotate GO IDs with GO Terms in the df_split output. 
# I tried splitting the GO.ID and GO.Term separately and then recombining them but realized the dimensions of the files were different. Part of the reason is that different databases may have annotated some of the GO Terms differently so removing duplicates did not work. For example,ATPase activity and ATP hydrolysis activity are technically the same. Thus, I am doing a 1:1 comparison of GO.ID and GO.Term against the GO.db.

goterms <- Term(GOTERM) # extract terms
ontology <- Ontology(GOTERM)
df <- as.data.frame(goterms, col_names = TRUE) # create a dataframe of goterms
df_ont <- as.data.frame(ontology, col_names = TRUE)
write.csv(df, file="goterms.csv", row.names = TRUE) # write as a csv
write.csv(df_ont, file="ontology.csv", row.names = TRUE) # write as a csv

go <- read_csv("goterms.csv", col_names = TRUE) # read the csv back into R to manipulate (I realize this could be a roundabout way to get the formatting correct but it worked!)
ont <- read_csv("ontology.csv", col_names = TRUE)
head(go)
colnames(go) <- c("GO.ID", "GO.Term") # rename col names
head(go)
dim(go) # 43559 
colnames(ont) <- c("GO.ID", "GO.Ontology") # rename col names
head(ont)
dim(ont) # 43559 

# head(df_split) # check col names of the split annotation dataframe
head(bind_df)
tail(bind_df)

universe <- left_join(bind_df, go, by="GO.ID") # join the two dataframes using GO.ID, left join includes all rows in df_split in this case
head(universe)
universe <- left_join(universe, ont, by = "GO.ID") # join dataframe with ontology
head(universe)
tail(universe)
dim(universe) # should be 30961
universe <- universe[, c(1, 3, 2, 4, 5)] # reorder columns 

## this file includes gene_id which is the unique gene identifier. this is necessary to do gene enrichment. The KO which includes KOs and then accession #s for non-KO annotated genes is necessary for heatmap visualizations.

write.csv(universe, file="tricho_GO_KO_database.csv", row.names = TRUE) # will serve as universe for heatmap visualizations
# I have a lotttttt of duplicated stuff
```
