---
title: "Tricho_Fewarming_DifferentalGeneExpression"
author: "Nina Yang, PhD"
adapted by: "Kyla Kelly, PhD Candidate, Hutchins Lab, USC"
date: "12/16/2022"
updated: "2/21/2022"
output: html_document
---

## Background

This document outlines the steps taken to for the Venn Diagram and clusterProfiler analyses.


**1. Install (if required) & upload packages into your current session**

```{r install-packages, eval = TRUE, message = FALSE, warning = FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE))
  install.packages("BiocManager")
if(!require(dplyr)){install.packages("dplyr")}
if(!require(tidyverse)){install.packages("tidyverse")}
if(!require(VennDiagram)){install.packages("VennDiagram")}
if(!require(VennDiagram)){install.packages("gridExtra")}
if(!require(VennDiagram)){install.packages("grid")}
if(!require(VennDiagram)){install.packages("lattice")}
```

**2. Upload the package into your current session**
```{r library, eval = TRUE, message = FALSE, warning = FALSE}
library("tidyverse")
library("clusterProfiler")
library("VennDiagram")
library("gridExtra")
library("grid")
library("ggplot2")
library("lattice")
```

# set up the BP, MF, and CC things
```{r}
bg <- read_csv("tricho_GO_KO_database.csv", col_names = TRUE)
# should these NA values for gene_id be the # thing I use earlier?
head(bg)
tail(bg)

# BP
bg_BP <- filter(bg, GO.Ontology == "BP")
head(bg_BP)
nrow(bg_BP)

go2gene_BP <- bg_BP[, c("GO.ID", "gene_id")]
go2name_BP <- bg_BP[, c("GO.ID", "GO.Term")]

# MF
bg_MF <- filter(bg, GO.Ontology == "MF")
head(bg_MF)
nrow(bg_MF)

go2gene_MF <- bg_MF[, c("GO.ID", "gene_id")]
go2name_MF <- bg_MF[, c("GO.ID", "GO.Term")]

# CC
bg_CC <- filter(bg, GO.Ontology == "CC")
head(bg_CC)
nrow(bg_CC)

go2gene_CC <- bg_CC[, c("GO.ID", "gene_id")]
go2name_CC <- bg_CC[, c("GO.ID", "GO.Term")]

# KO
bg_KO <- read_csv("filtered_tricho_kofamscan.csv", col_names = TRUE) # this is the file that will be used for the universe or background genes
colnames(bg_KO)
bg_KO <- bg_KO$gene_id
head(bg_KO)
tail(bg_KO)
```

**3. Process the DESeq2 outputs to generate Venn Diagrams and clusterProfiler inputs**
## deleted the day vs night stuff, tho the way the comparisons are laid out might help me later 

### 3.1 Replete vs Limited (Day)

```{r RL_input, eval = TRUE, message = FALSE, warning = FALSE}

cts <- read_csv("../deseq2/tricho_trimmed_forDESeq2.csv")
head(cts) 
gene_id <- cts[c(1:2)] # extract gene_id to add to the other dataframes because this number was left off. The numeric ordering of these files is the same as the input file which has gene_id as well.
head(gene_id)

# set up replete vs limited dataframes (day)
## Replete v Limited Replete means that positive is upregulated for Replete relative to Limited
sub <- read_csv("../deseq2/outputs/R22vL22_DESeqresults.csv")
head(sub)

opt <- read_csv("../deseq2/outputs/R27vL27_DESeqresults.csv")
head(opt)

supra <- read_csv("../deseq2/outputs/R32vL32_DESeqresults.csv")
head(supra)

# sub 
sub_df <- cbind(gene_id, sub) # append gene_ids
summary(sub_df)
head(sub_df)
subRL <- dplyr::select(sub_df, gene_id, log2FoldChange, padj) # select relevant columns
head(subRL)
dim(subRL) # 4475 all DGEs
subRL.vector <- dplyr::filter(subRL, padj < 0.05) # all the differentially expressed sub RL genes 
head(subRL.vector)
subRL.vector <-  dplyr::select(subRL.vector, gene_id, log2FoldChange) #remove padj column
dim(subRL.vector) # 668 significant DGEs
head(subRL.vector)
names(subRL.vector)[2] <- "subRL_FC"
head(subRL.vector)

# opt 
opt_df <- cbind(gene_id, opt) # append gene_ids
summary(opt_df)
head(opt_df)
optRL <- dplyr::select(opt_df, gene_id, log2FoldChange, padj) # select relevant columns
head(optRL)
dim(optRL) #4475
optRL.vector <- dplyr::filter(optRL, padj < 0.05) # all the differentially expressed opt RL genes 
head(optRL.vector)
optRL.vector <-  dplyr::select(optRL.vector, gene_id, log2FoldChange) #remove padj column
dim(optRL.vector) #1676
head(optRL.vector)
names(optRL.vector)[2] <- "optRL_FC"
head(optRL.vector)

# supra 
supra_df <- cbind(gene_id, supra) # append gene_ids
summary(supra_df)
head(supra_df)
supraRL <- dplyr::select(supra_df, gene_id, log2FoldChange, padj) # select relevant columns
head(supraRL)
dim(supraRL) #4475
supraRL.vector <- dplyr::filter(supraRL, padj < 0.05) # all the differentially expressed supra RL genes 
head(supraRL.vector)
supraRL.vector <-  dplyr::select(supraRL.vector, gene_id, log2FoldChange) #remove padj column
dim(supraRL.vector) #728
head(supraRL.vector)
names(supraRL.vector)[2] <- "supraRL_FC"
head(supraRL.vector)

# build a replete vs limited dataframe that combines, sub, opt, and supra differentially expressed genes

# start with the list of all relevant genes (the full list of deduped and trimmed list of genes)

RL.vector <- left_join(gene_id, subRL.vector, by = "gene_id")
head(RL.vector)
dim(RL.vector) #4475 x 3

RL.vector <- left_join(RL.vector, optRL.vector, by = "gene_id")
dim(RL.vector) #4475 x 4
head(RL.vector)

RL.vector <- left_join(RL.vector, supraRL.vector, by = "gene_id")
head(RL.vector)
dim(RL.vector) #4475 x 5

head(RL.vector, 10)
tail(RL.vector, 10)

# so now I have the log fold change values for each gene_id for repelte v limited for each temp
```

# venn diagram
```{r RL_venn, eval = TRUE, message = FALSE, warning = FALSE}

# build Venn Diagram input
venn_df <- RL.vector %>%
  gather(key = exp_type, value = foldchange, -c(gene_id))
glimpse(venn_df)

#### this is for all diel genes
venn_sub <- venn_df %>% 
  filter(exp_type == "subRL_FC" & !is.na(foldchange)) %>% 
  pull(gene_id)

venn_opt <- venn_df %>% 
  filter(exp_type == "optRL_FC" & !is.na(foldchange)) %>% 
  pull(gene_id)

venn_supra <- venn_df %>% 
  filter(exp_type == "supraRL_FC" & !is.na(foldchange)) %>% 
  pull(gene_id)

## this is for positive fold change (increase in abundance)
pos_venn_sub <- venn_df %>% 
  filter(exp_type == "subRL_FC" & !is.na(foldchange) & foldchange > 0) %>% 
  pull(gene_id)

pos_venn_opt <- venn_df %>% 
  filter(exp_type == "optRL_FC" & !is.na(foldchange) & foldchange > 0) %>% 
  pull(gene_id)

pos_venn_supra <- venn_df %>% 
  filter(exp_type == "supraRL_FC" & !is.na(foldchange) & foldchange > 0) %>% 
  pull(gene_id)

## this is for negative fold change (decrease in abundance)

neg_venn_sub <- venn_df %>% 
  filter(exp_type == "subRL_FC" & !is.na(foldchange) & foldchange < 0) %>% 
  pull(gene_id)

neg_venn_opt <- venn_df %>% 
  filter(exp_type == "optRL_FC" & !is.na(foldchange) & foldchange < 0) %>% 
  pull(gene_id)

neg_venn_supra <- venn_df %>% 
  filter(exp_type == "supraRL_FC" & !is.na(foldchange) & foldchange < 0) %>% 
  pull(gene_id)

summary(pos_venn_sub) #341
summary(pos_venn_opt) #917
summary(pos_venn_supra) #414

summary(neg_venn_sub) #327
summary(neg_venn_opt) #759
summary(neg_venn_supra) #314

venn_list_pos <- list("22C" = pos_venn_sub,
                      "27C" = pos_venn_opt,
                      "32C" = pos_venn_supra)

venn_list_neg <- list("22C" = neg_venn_sub,
                      "27C" = neg_venn_opt,
                      "32C" = neg_venn_supra)


# Upregulated RL DEG Venn Diagrams - build Venn Diagram and extract gene lists
# ven_list_pos = genes upregulated in replete v. limited, meaning upregulated genes are upregulated for replete, relative to limited
# or downregulated for limited, relative to replete

venn <- venn.diagram(venn_list_pos, NULL,
                          col = "transparent",
                          fill = c("#350498", "#B93289", "#FBA139"),
                          alpha = 0.50,
                          cex = 0.8,
                          fontfamily = "sans",
                          fontface = "bold",
                          cat.col = c("#350498", "#B93289", "#FBA139"),
                          cat.cex = 0.8,
                          cat.fontfamily = "sans",
                          margin = 0.2,
                          main.fontfamily = "sans",
                          print.mode = c("raw","percent"), # Show both numbers and percent
                          main.pos = c(0.5,0.9))


grid.draw(venn)

# this creates a file that is upregulated for Fe-replete which is downregulated for Fe-limited
# hmm im not sure because it says the same thing below
ggsave("vd_Fe/limited_down.pdf", plot = venn, height=120, width=120, units='mm', dpi=500)

dev.off()

## extract the overlap analysis so you have the list of genes that are in each section of the venn diagram

list <- calculate.overlap(venn_list_pos)

summary(list)

# core genes shared across day treatments
write.csv(list$a5,"vd_Fe/Limited_Core_down.csv", row.names = TRUE)

# shared between 22 & 27
write.csv(list$a2,"vd_Fe/Limited_22_27_Shared_down.csv", row.names = TRUE)

# shared between 22 & 32
write.csv(list$a4,"vd_Fe/Limited_22_32_Shared_down.csv", row.names = TRUE)

# shared between 27 & 32
write.csv(list$a6,"vd_Fe/Limited_27_32_Shared_down.csv", row.names = TRUE)

# unique 22C
write.csv(list$a1,"vd_Fe/Limited_22_Unique_down.csv", row.names = TRUE)

# unique 27C
write.csv(list$a3,"vd_Fe/Limited_27_Unique_down.csv", row.names = TRUE)

# unique 32C
write.csv(list$a7,"vd_Fe/Limited_32_Unique_down.csv", row.names = TRUE)


  # Downregulated RL DEG Venn Diagrams - build Venn Diagram and extract gene lists (upregulated for Fe-limited)

venn <- venn.diagram(venn_list_neg, NULL,
                          col = "transparent",
                          fill = c("#350498", "#B93289", "#FBA139"),
                          alpha = 0.50,
                          cex = 0.8,
                          fontfamily = "sans",
                          fontface = "bold",
                          cat.col = c("#350498", "#B93289", "#FBA139"),
                          cat.cex = 0.8,
                          cat.fontfamily = "sans",
                          margin = 0.2,
                          main.fontfamily = "sans",
                          print.mode = c("raw","percent"), # Show both numbers and percent
                          main.pos = c(0.5,0.9))

grid.draw(venn)

# this creates a file that is upregulated for Fe-replete which is downregulated for Fe-limited
# hmm im not sure because it says the same thing above
ggsave("vd_Fe/Limited_up.pdf", plot = venn, height=120, width=120, units='mm', dpi=500)

dev.off()


## extract the overlap analysis so you have the list of genes that are in each section of the venn diagram

list <- calculate.overlap(venn_list_neg)

summary(list)

# core genes shared across day treatments
write.csv(list$a5,"vd_Fe/Limited_Core_up.csv", row.names = TRUE)

# shared between 22 & 27
write.csv(list$a2,"vd_Fe/Limited_22_27_Shared_up.csv", row.names = TRUE)

# shared between 22 & 32
write.csv(list$a4,"vd_Fe/Limited_22_32_Shared_up.csv", row.names = TRUE)

# shared between 27 & 32
write.csv(list$a6,"vd_Fe/Limited_27_32_Shared_up.csv", row.names = TRUE)

# unique 22C
write.csv(list$a1,"vd_Fe/Limited_22_Unique_up.csv", row.names = TRUE)

# unique 27C
write.csv(list$a3,"vd_Fe/Limited_27_Unique_up.csv", row.names = TRUE)

# unique 32C
write.csv(list$a7,"vd_Fe/Limited_32_Unique_up.csv", row.names = TRUE)

```


# cluster profiler core limited down and up
```{r Limited_down/up_core_clusterProfiler, eval = TRUE, message = FALSE, warning = FALSE}

## down 

df <- read_csv("vd_Fe/Limited_Core_down.csv")
head(df)
dim(df) #107
df_ids <- df$x
dim(df) #107

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich

  write.csv(enrich, file="../cluster_profiler/Fe_RL/Core_down_enrich_BP.csv", row.names = TRUE) # enricher results

# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Core_down_enrich_MF.csv", row.names = TRUE) # enricher results


# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_CC, TERM2NAME = go2name_CC, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Core_down_enrich_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Core_down_enrich_KO.csv", row.names = TRUE) # enricher results


## up

df <- read_csv("vd_Fe/Limited_Core_up.csv")
head(df)
dim(df) #83
df_ids <- df$x
dim(df)

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Core_up_enrich_BP.csv", row.names = TRUE) # enricher results

# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Core_up_enrich_MF.csv", row.names = TRUE) # enricher results


# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_CC, TERM2NAME = go2name_CC, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Core_up_enrich_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Core_up_enrich_KO.csv", row.names = TRUE) # enricher results

```

# cluster profiler shared/unique limited up/down
```{r Limited_up/down_shared_unique_clusterProfiler, eval = TRUE, message = FALSE, warning = FALSE}

# Shared 22 and 27

df <- read_csv("vd_Fe/Limited_22_27_Shared_up.csv")

head(df)
df_ids <- df$x

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Shared_22_27_up_enrich_BP.csv", row.names = TRUE) # enricher results


# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Shared_22_27_up_enrich_MF.csv", row.names = TRUE) # enricher results


# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_CC, TERM2NAME = go2name_CC, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Shared_22_27_up_enrich_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Shared_22_27_up_enrich_KO.csv", row.names = TRUE) # enricher results


# Shared 22 and 32

df <- read_csv("vd_Fe/Limited_22_32_Shared_up.csv")
head(df)
df_ids <- df$x

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Shared_22_32_up_enrich_BP.csv", row.names = TRUE) # enricher results

# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Shared_22_32_up_enrich_MF.csv", row.names = TRUE) # enricher results


# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_CC, TERM2NAME = go2name_CC, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Shared_22_32_up_enrich_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Shared_22_32_up_enrich_KO.csv", row.names = TRUE) # enricher results


# shared 27 and 32

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Shared_27_32_up_enrich_BP.csv", row.names = TRUE) # enricher results

# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Shared_27_32_up_enrich_MF.csv", row.names = TRUE) # enricher results


# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_CC, TERM2NAME = go2name_CC, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Shared_27_32_up_enrich_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Shared_27_32_up_enrich_KO.csv", row.names = TRUE) # enricher results

# Unique 22

df <- read_csv("vd_Fe/Limited_22_Unique_up.csv")
head(df)
df_ids <- df$x
dim(df)

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Unique_22_up_enrich_BP.csv", row.names = TRUE) # enricher results

# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Unique_22_up_enrich_MF.csv", row.names = TRUE) # enricher results


# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_CC, TERM2NAME = go2name_CC, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Unique_22_up_enrich_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Unique_22_up_enrich_KO.csv", row.names = TRUE) # enricher results


# Unique 27

df <- read_csv("vd_Fe/Limited_27_Unique_up.csv")
head(df)
df_ids <- df$x
dim(df)

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Unique_27_up_enrich_BP.csv", row.names = TRUE) # enricher results


# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Unique_27_up_enrich_MF.csv", row.names = TRUE) # enricher results


# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_CC, TERM2NAME = go2name_CC, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Unique_27_up_enrich_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Unique_27_up_enrich_KO.csv", row.names = TRUE) # enricher results


# Unique 32
df <- read_csv("vd_Fe/Limited_32_Unique_up.csv")
head(df)
df_ids <- df$x
dim(df)

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Unique_32_up_enrich_BP.csv", row.names = TRUE) # enricher results


# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich


write.csv(enrich, file="../cluster_profiler/Fe_RL/Unique_32_up_enrich_MF.csv", row.names = TRUE) # enricher results


# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_CC, TERM2NAME = go2name_CC, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Unique_32_up_enrich_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Unique_32_up_enrich_KO.csv", row.names = TRUE) # enricher results


# Shared 22 and 27

df <- read_csv("vd_Fe/Limited_22_27_Shared_down.csv")

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Shared_22_27_down_enrich_BP.csv", row.names = TRUE) # enricher results


# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Shared_22_27_down_enrich_MF.csv", row.names = TRUE) # enricher results


# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_CC, TERM2NAME = go2name_CC, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Shared_22_27_down_enrich_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Shared_22_27_down_enrich_KO.csv", row.names = TRUE) # enricher results


# Shared 22 and 32

df <- read_csv("vd_Fe/Limited_22_32_Shared_down.csv")
head(df)
df_ids <- df$x

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Shared_22_32_down_enrich_BP.csv", row.names = TRUE) # enricher results

# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Shared_22_32_down_enrich_MF.csv", row.names = TRUE) # enricher results


# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_CC, TERM2NAME = go2name_CC, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Shared_22_32_down_enrich_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Shared_22_32_down_enrich_KO.csv", row.names = TRUE) # enricher results


# shared 27 and 32

df <- read_csv("vd_Fe/Limited_27_32_Shared_down.csv")
head(df)
df_ids <- df$x

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Shared_27_32_down_enrich_BP.csv", row.names = TRUE) # enricher results

# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Shared_27_32_down_enrich_MF.csv", row.names = TRUE) # enricher results


# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_CC, TERM2NAME = go2name_CC, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Shared_27_32_down_enrich_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Shared_27_32_down_enrich_KO.csv", row.names = TRUE) # enricher results

# Unique 22

df <- read_csv("vd_Fe/Limited_22_Unique_down.csv")
head(df)
df_ids <- df$x
dim(df)

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Unique_22_down_enrich_BP.csv", row.names = TRUE) # enricher results

# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Unique_22_down_enrich_MF.csv", row.names = TRUE) # enricher results


# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_CC, TERM2NAME = go2name_CC, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Unique_22_down_enrich_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Unique_22_down_enrich_KO.csv", row.names = TRUE) # enricher results


# Unique 27

df <- read_csv("vd_Fe/Limited_27_Unique_down.csv")
head(df)
df_ids <- df$x
dim(df)

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Unique_27_down_enrich_BP.csv", row.names = TRUE) # enricher results


# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Unique_27_down_enrich_MF.csv", row.names = TRUE) # enricher results


# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Unique_27_down_enrich_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Unique_27_down_enrich_KO.csv", row.names = TRUE) # enricher results


# Unique 32

df <- read_csv("vd_Fe/Limited_32_Unique_down.csv")
head(df)
df_ids <- df$x
dim(df)

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Unique_32_down_enrich_BP.csv", row.names = TRUE) # enricher results


# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Unique_32_down_enrich_MF.csv", row.names = TRUE) # enricher results


# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_CC, TERM2NAME = go2name_CC, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Unique_32_down_enrich_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/Fe_RL/Unique_32_down_enrich_KO.csv", row.names = TRUE) # enricher results

```

### 3.2: R vs L for different temperatures

```{r RL_input_temp, eval = TRUE, message = FALSE, warning = FALSE}

cts <- read_csv("../deseq2/tricho_trimmed_forDESeq2.csv")
head(cts) 

gene_id <- cts[c(1:2)] # extract gene_id to add to the other dataframes because this number was left off. The numeric ordering of these files is the same as the input file which has gene_id as well.
head(gene_id)

# set up replete vs limited dataframes (day)
sub <- read_csv("../deseq2/outputs/R22vL22_DESeqresults.csv")
head(sub)

opt <- read_csv("../deseq2/outputs/R27vL27_DESeqresults.csv")
head(opt)

supra <- read_csv("../deseq2/outputs/R32vL32_DESeqresults.csv")
head(supra)

# sub 
sub_df <- cbind(gene_id, sub) # append gene_ids
summary(sub_df)
head(sub_df)
subRL <- dplyr::select(sub_df, gene_id, log2FoldChange, padj) # select relevant columns
head(subRL)
dim(subRL) #4475
subRL.vector <- dplyr::filter(subRL, padj < 0.05) # all the differentially expressed sub RL genes 
head(subRL.vector)
subRL.vector <- dplyr::select(subRL.vector, gene_id, log2FoldChange) #remove padj column
dim(subRL.vector) #668
head(subRL.vector)
names(subRL.vector)[2] <- "subRL_FC"
head(subRL.vector)
tail(subRL.vector)

# for replete
sub_rep <- filter(subRL.vector, subRL_FC > 0 )
dim(sub_rep) #341

# for limited
sub_lim <- filter(subRL.vector, subRL_FC < 0 )
dim(sub_lim) #327

# opt 
opt_df <- cbind(gene_id, opt) # append gene_ids
summary(opt_df)
head(opt_df)
optRL <- dplyr::select(opt_df, gene_id, log2FoldChange, padj) # select relevant columns
head(optRL)
dim(optRL) #4475
optRL.vector <- dplyr::filter(optRL, padj < 0.05) # all the differentially expressed opt RL genes 
head(optRL.vector)
optRL.vector <- dplyr::select(optRL.vector, gene_id, log2FoldChange) #remove padj column
dim(optRL.vector) #1676
head(optRL.vector)
names(optRL.vector)[2] <- "optRL_FC"
head(optRL.vector)

# for replete
opt_rep <- filter(optRL.vector, optRL_FC > 0 )
dim(opt_rep) #917

# for limited
opt_lim <- filter(optRL.vector, optRL_FC < 0 )
dim(opt_lim) #759

# supra 
supra_df <- cbind(gene_id, supra) # append gene_ids
summary(supra_df)
head(supra_df)
supraRL <- dplyr::select(supra_df, gene_id, log2FoldChange, padj) # select relevant columns
head(supraRL)
dim(supraRL) #728
supraRL.vector <- dplyr::filter(supraRL, padj < 0.05) # all the differentially expressed supra RL genes 
head(supraRL.vector)
supraRL.vector <- dplyr::select(supraRL.vector, gene_id, log2FoldChange) #remove padj column
dim(supraRL.vector) #728
head(supraRL.vector)
names(supraRL.vector)[2] <- "supraRL_FC"
head(supraRL.vector)

# for replete
supra_rep <- filter(supraRL.vector, supraRL_FC > 0 )
dim(supra_rep) #414

# for limited
supra_lim <- filter(supraRL.vector, supraRL_FC < 0 )
dim(supra_lim) #314

# need to run clusterprofilers on these dataframes
head(sub_rep)
head(sub_lim)
head(opt_rep)
head(opt_lim)
head(supra_rep)
head(supra_lim)
```

# cluster profiler 
```{r RL_input_temp, eval = TRUE, message = FALSE, warning = FALSE}

## Sub rep
df <- sub_rep
head(df)
df_ids <- df$gene_id
head(df_ids)

# BP
head(df)
df_ids <- df$gene_id
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich
write.csv(enrich, file="../cluster_profiler/temp/sub_L_down_BP.csv", row.names = TRUE) # enricher results

# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich
write.csv(enrich, file="../cluster_profiler/temp/sub_L_down_MF.csv", row.names = TRUE) # enricher results

# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_CC, TERM2NAME = go2name_CC, minGSSize = 5)
enrich
write.csv(enrich, file="../cluster_profiler/temp/sub_L_down_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich
write.csv(enrich, file="../cluster_profiler/temp/sub_L_down_KO.csv", row.names = TRUE) # enricher results

## sub lim 
df <- sub_lim
head(df)
df_ids <- df$gene_id

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich
write.csv(enrich, file="../cluster_profiler/temp/sub_L_up_BP.csv", row.names = TRUE) # enricher results

# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich
write.csv(enrich, file="../cluster_profiler/temp/sub_L_up_MF.csv", row.names = TRUE) # enricher results

# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_CC, TERM2NAME = go2name_CC, minGSSize = 5)
enrich
write.csv(enrich, file="../cluster_profiler/temp/sub_L_up_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich
write.csv(enrich, file="../cluster_profiler/temp/sub_L_up_KO.csv", row.names = TRUE) # enricher results

## opt rep
df <- opt_rep
head(df)
df_ids <- df$gene_id

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich
write.csv(enrich, file="../cluster_profiler/temp/opt_L_down_BP.csv", row.names = TRUE) # enricher results

# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich
write.csv(enrich, file="../cluster_profiler/temp/opt_L_down_MF.csv", row.names = TRUE) # enricher results

# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_CC, TERM2NAME = go2name_CC, minGSSize = 5)
enrich
write.csv(enrich, file="../cluster_profiler/temp/opt_L_down_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich
write.csv(enrich, file="../cluster_profiler/temp/opt_L_down_KO.csv", row.names = TRUE) # enricher results

## opt lim 
df <- opt_lim
head(df)
df_ids <- df$gene_id

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich
write.csv(enrich, file="../cluster_profiler/temp/opt_L_up_BP.csv", row.names = TRUE) # enricher results

# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich
write.csv(enrich, file="../cluster_profiler/temp/opt_L_up_MF.csv", row.names = TRUE) # enricher results

# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_CC, TERM2NAME = go2name_CC, minGSSize = 5)
enrich
write.csv(enrich, file="../cluster_profiler/temp/opt_L_up_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich
write.csv(enrich, file="../cluster_profiler/temp/opt_L_up_KO.csv", row.names = TRUE) # enricher results

## supra rep
df <- supra_rep
head(df)
df_ids <- df$gene_id

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich
write.csv(enrich, file="../cluster_profiler/temp/supra_L_down_BP.csv", row.names = TRUE) # enricher results

# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich
write.csv(enrich, file="../cluster_profiler/temp/supra_L_down_MF.csv", row.names = TRUE) # enricher results

# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_CC, TERM2NAME = go2name_CC, minGSSize = 5)
enrich
write.csv(enrich, file="../cluster_profiler/temp/supra_L_down_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich
write.csv(enrich, file="../cluster_profiler/temp/supra_L_down_KO.csv", row.names = TRUE) # enricher results

## supra lim 
df <- supra_lim
head(df)
df_ids <- df$gene_id

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich
write.csv(enrich, file="../cluster_profiler/temp/supra_L_up_BP.csv", row.names = TRUE) # enricher results

# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich
write.csv(enrich, file="../cluster_profiler/temp/supra_L_up_MF.csv", row.names = TRUE) # enricher results

# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_CC, TERM2NAME = go2name_CC, minGSSize = 5)
enrich
write.csv(enrich, file="../cluster_profiler/temp/supra_L_up_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich
write.csv(enrich, file="../cluster_profiler/temp/supra_L_up_KO.csv", row.names = TRUE) # enricher results
```


### 3.3 Temperatures (22 v 27, 27 v 32)
```{r day_temp, eval = TRUE, message = FALSE, warning = FALSE}

cts <- read_csv("../deseq2/tricho_trimmed_forDESeq2.csv")
head(cts) 

gene_id <- cts[c(1:2)] # extract gene_id to add to the other dataframes because this number was left off. The numeric ordering of these files is the same as the input file which has gene_id as well.
head(gene_id)

# set up replete day dataframes

## R22 v R27 means that positive is upregulated for R22 relative to R27
sub <- read_csv("../deseq2/outputs/R22vR27_DESeqresults.csv")
head(sub)

## R27 v R32 means that positive is upregulated for R22 relative to R32
opt <- read_csv("../deseq2/outputs/R27vR32_DESeqresults.csv")
head(opt)

## R22 v R32 means that positive is upregulated for R22 relative to R32
supra <- read_csv("../deseq2/outputs/R22vR32_DESeqresults.csv")
head(supra)

# sub 
sub_df <- cbind(gene_id, sub) # append gene_ids
summary(sub_df)
head(sub_df)
subR <- dplyr::select(sub_df, gene_id, log2FoldChange, padj) # select relevant columns
head(subR)
dim(subR) #4475
subR.vector <- dplyr::filter(subR, padj < 0.05) 
head(subR.vector)
subR.vector <- dplyr::select(subR.vector, gene_id, log2FoldChange) #remove padj column
dim(subR.vector) #922
head(subR.vector)
names(subR.vector)[2] <- "subR_FC"
head(subR.vector)
subR.vector.up <- dplyr::filter(subR.vector, subR_FC > 0) # upregulated genes
subR.vector.down <- dplyr::filter(subR.vector, subR_FC < 0) # downregulated genes

write.csv(subR.vector.up, file="../cluster_profiler/vectors/R22vR27_up_vector.csv", row.names = TRUE) # for clusterProfiler

write.csv(subR.vector.down, file="../cluster_profiler/vectors/R22vR27_down_vector.csv", row.names = TRUE) # for clusterProfiler

# opt 
opt_df <- cbind(gene_id, opt) # append gene_ids
summary(opt_df)
head(opt_df)
optR <- dplyr::select(opt_df, gene_id, log2FoldChange, padj) # select relevant columns
head(optR)
dim(optR) #4475
optR.vector <- dplyr::filter(optR, padj < 0.05) # all the differentially expressed opt RL genes 
head(optR.vector)
optR.vector <- dplyr::select(optR.vector, gene_id, log2FoldChange) #remove padj column
dim(optR.vector) #539
head(optR.vector)
names(optR.vector)[2] <- "optR_FC"
head(optR.vector)
optR.vector.up <- dplyr::filter(optR.vector, optR_FC > 0) 
optR.vector.down <- dplyr::filter(optR.vector, optR_FC < 0) 

write.csv(optR.vector.up, file="../cluster_profiler/vectors/R27vR32_up_vector.csv", row.names = TRUE) # for clusterProfiler

write.csv(optR.vector.down, file="../cluster_profiler/vectors/R27vR32_down_vector.csv", row.names = TRUE) # for clusterProfiler

# supra 
supra_df <- cbind(gene_id, supra) # append gene_ids
summary(supra_df)
head(supra_df)
supraR <- dplyr::select(supra_df, gene_id, log2FoldChange, padj) # select relevant columns
head(supraR)
dim(supraR) #4475
supraR.vector <- dplyr::filter(supraR, padj < 0.05)
head(supraR.vector)
supraR.vector <- dplyr::select(supraR.vector, gene_id, log2FoldChange) #remove padj column
dim(supraR.vector)#1175
head(supraR.vector)
names(supraR.vector)[2] <- "supraR_FC"
head(supraR.vector)
supraR.vector.up <- dplyr::filter(supraR.vector, supraR_FC > 0)
supraR.vector.down <- dplyr::filter(supraR.vector, supraR_FC < 0)

write.csv(supraR.vector.up, file="../cluster_profiler/vectors/R22vR32_up_vector.csv", row.names = TRUE) # for clusterProfiler

write.csv(supraR.vector.down, file="../cluster_profiler/vectors/R22vR32_down_vector.csv", row.names = TRUE) # for clusterProfiler


# set up limited

## L22 v L27 means that positive is upregulated for L22 relative to L27
sub <- read_csv("../deseq2/outputs/L22vL27_DESeqresults.csv")
head(sub)

## L27 v L32 means that positive is upregulated for L27 relative to L32
opt <- read_csv("../deseq2/outputs/L27vL32_DESeqresults.csv")
head(opt)

## L22 v L32 means that positive is upregulated for L22 relative to L32
supra <- read_csv("../deseq2/outputs/L22vL32_DESeqresults.csv")
head(supra)

# sub 
sub_df <- cbind(gene_id, sub) # append gene_ids
summary(sub_df)
head(sub_df)
subL <- dplyr::select(sub_df, gene_id, log2FoldChange, padj) # select relevant columns
head(subL)
dim(subL) #4475
subL.vector <- dplyr::filter(subL, padj < 0.05)
head(subL.vector) #1224
subL.vector <- dplyr::select(subL.vector, gene_id, log2FoldChange) #remove padj column
dim(subL.vector)
head(subL.vector)
names(subL.vector)[2] <- "subL_FC"
head(subL.vector)
subL.vector.up <- dplyr::filter(subL.vector, subL_FC > 0)
subL.vector.down <- dplyr::filter(subL.vector, subL_FC < 0)

write.csv(subL.vector.up, file="../cluster_profiler/vectors/L22vL27_up_vector.csv", row.names = TRUE) # for clusterProfiler

write.csv(subL.vector.down, file="../cluster_profiler/vectors/L22vL27_down_vector.csv", row.names = TRUE) # for clusterProfiler

# opt 
opt_df <- cbind(gene_id, opt) # append gene_ids
summary(opt_df)
head(opt_df)
optL <- dplyr::select(opt_df, gene_id, log2FoldChange, padj) # select relevant columns
head(optL)
dim(optL) #4475
optL.vector <- dplyr::filter(optL, padj < 0.05) # all the differentially expressed opt RL genes 
head(optL.vector)
optL.vector <- dplyr::select(optL.vector, gene_id, log2FoldChange) #remove padj column
dim(optL.vector) #1265
head(optL.vector)
names(optL.vector)[2] <- "optL_FC"
head(optL.vector)
optL.vector.up <- dplyr::filter(optL.vector, optL_FC > 0)  
optL.vector.down <- dplyr::filter(optL.vector, optL_FC < 0) 

write.csv(optL.vector.up, file="../cluster_profiler/vectors/L27vL32_up_vector.csv", row.names = TRUE) # for clusterProfiler

write.csv(optL.vector.down, file="../cluster_profiler/vectors/L27vL32_down_vector.csv", row.names = TRUE) # for clusterProfiler

# supra 
supra_df <- cbind(gene_id, supra) # append gene_ids
summary(supra_df)
head(supra_df)
supraL <- dplyr::select(supra_df, gene_id, log2FoldChange, padj) # select relevant columns
head(supraL)
dim(supraL) #4475
supraL.vector <- dplyr::filter(supraL, padj < 0.05) # all the differentially expressed supra RL genes 
head(supraL.vector)
supraL.vector <- dplyr::select(supraL.vector, gene_id, log2FoldChange) #remove padj column
dim(supraL.vector) #1333
head(supraL.vector)
names(supraL.vector)[2] <- "supraL_FC"
head(supraL.vector)
supraL.vector.up <- dplyr::filter(supraL.vector, supraL_FC > 0)
supraL.vector.down <- dplyr::filter(supraL.vector, supraL_FC < 0) 

write.csv(supraL.vector.up, file="../cluster_profiler/vectors/L22vL32_up_vector.csv", row.names = TRUE) # for clusterProfiler

write.csv(supraL.vector.down, file="../cluster_profiler/vectors/L22vL32_down_vector.csv", row.names = TRUE) # for clusterProfiler
```

# cluster profiler for temp v temp within each nutrient treatment 
```{r 22RDv27RD_clusterProfiler, eval = TRUE, message = FALSE, warning = FALSE}


#### cluster profiler replete 22 v 27
# 22R vs 27R
# means that positive is upregulated for replete, relative to limited

## UP
df <- read_csv("../cluster_profiler/vectors/R22vR27_up_vector.csv")
head(df)
df_ids <- df$gene_id

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/R_temp_comparisons/R22vR27_up_BP.csv", row.names = TRUE) # enricher results

# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/R_temp_comparisons/R22vR27_up_MF.csv", row.names = TRUE) # enricher results

# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_CC, TERM2NAME = go2name_CC, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/R_temp_comparisons/R22vR27_up_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/R_temp_comparisons/R22vR27_up_KO.csv", row.names = TRUE) # enricher results

## DOWN
df <- read_csv("../cluster_profiler/vectors/R22vR27_down_vector.csv")
head(df)
df_ids <- df$gene_id

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich
write.csv(enrich, file="../cluster_profiler/R_temp_comparisons/R22vR27_down_BP.csv", row.names = TRUE) # enricher results

# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/R_temp_comparisons/R22vR27_down_MF.csv", row.names = TRUE) # enricher results

# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_CC, TERM2NAME = go2name_CC, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/R_temp_comparisons/R22vR27_down_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/R_temp_comparisons/R22vR27_down_KO.csv", row.names = TRUE) # enricher results


#### cluster profiler 27 v 32 replete
# 27R vs 32R

## UP
df <- read_csv("../cluster_profiler/vectors/R27vR32_up_vector.csv")
head(df)
df_ids <- df$gene_id

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/R_temp_comparisons/R27vR32_up_BP.csv", row.names = TRUE) # enricher results

# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/R_temp_comparisons/R27vR32_up_MF.csv", row.names = TRUE) # enricher results

# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_CC, TERM2NAME = go2name_CC, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/R_temp_comparisons/R27vR32_up_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/R_temp_comparisons/R27vR32_up_KO.csv", row.names = TRUE) # enricher results

## DOWN 
df <- read_csv("../cluster_profiler/vectors/R27vR32_down_vector.csv")
head(df)
df_ids <- df$gene_id

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/R_temp_comparisons/R27vR32_down_BP.csv", row.names = TRUE) # enricher results

# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/R_temp_comparisons/R27vR32_down_MF.csv", row.names = TRUE) # enricher results

# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_CC, TERM2NAME = go2name_CC, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/R_temp_comparisons/R27vR32_down_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/R_temp_comparisons/R27vR32_down_KO.csv", row.names = TRUE) # enricher results



#### cluster profiler 22 v 27 limited
# 22L vs 27L

## UP
df <- read_csv("../cluster_profiler/vectors/L22vL27_up_vector.csv")
head(df)
df_ids <- df$gene_id

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/L_temp_comparisons/L22vL27_up_BP.csv", row.names = TRUE) # enricher results

# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/L_temp_comparisons/L22vL27_up_MF.csv", row.names = TRUE) # enricher results

# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich
write.csv(enrich, file="../cluster_profiler/L_temp_comparisons/L22vL27_up_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/L_temp_comparisons/L22vL27_up_KO.csv", row.names = TRUE) # enricher results

## DOWN

df <- read_csv("../cluster_profiler/vectors/L22vL27_down_vector.csv")
head(df)
df_ids <- df$gene_id

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/L_temp_comparisons/L22vL27_down_BP.csv", row.names = TRUE) # enricher results

# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/L_temp_comparisons/L22vL27_down_MF.csv", row.names = TRUE) # enricher results

# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_CC, TERM2NAME = go2name_CC, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/L_temp_comparisons/L22vL27_down_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/L_temp_comparisons/L22vL27_down_KO.csv", row.names = TRUE) # enricher results

#### Cluster profiler 27 v 32 limited 

# 27L v 32L

## UP
df <- read_csv("../cluster_profiler/vectors/L27vL32_up_vector.csv")
head(df)
df_ids <- df$gene_id

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/L_temp_comparisons/L27vL32_up_BP.csv", row.names = TRUE) # enricher results

# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/L_temp_comparisons/L27vL32_up_MF.csv", row.names = TRUE) # enricher results

# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_CC, TERM2NAME = go2name_CC, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/L_temp_comparisons/L27vL32_up_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/L_temp_comparisons/L27vL32_up_KO.csv", row.names = TRUE) # enricher results


## DOWN
df <- read_csv("../cluster_profiler/vectors/L27vL32_down_vector.csv")
head(df)
df_ids <- df$gene_id

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/L_temp_comparisons/L27vL32_down_BP.csv", row.names = TRUE) # enricher results

# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/L_temp_comparisons/L27vL32_down_MF.csv", row.names = TRUE) # enricher results

# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_CC, TERM2NAME = go2name_CC, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/L_temp_comparisons/L27vL32_down_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/L_temp_comparisons/L27vL32_down_KO.csv", row.names = TRUE) # enricher results
```


# venn diagrams replete 
```{r R_venn, eval = TRUE, message = FALSE, warning = FALSE}

# set up replete day dataframes
sub <- read_csv("../deseq2/outputs/R22vR27_DESeqresults.csv")
head(sub)
supra <- read_csv("../deseq2/outputs/R27vR32_DESeqresults.csv")
head(supra)

# sub 
sub_df <- cbind(gene_id, sub) # append gene_ids
summary(sub_df)
head(sub_df)
subR <- dplyr::select(sub_df, gene_id, log2FoldChange, padj) # select relevant columns
head(subR)
dim(subR) #4475
subR.vector <- dplyr::filter(subR, padj < 0.05) 
head(subR.vector)
subR.vector <- dplyr::select(subR.vector, gene_id, log2FoldChange) #remove padj column
dim(subR.vector) #922
head(subR.vector)
names(subR.vector)[2] <- "subR_FC"
head(subR.vector)

# supra
supra_df <- cbind(gene_id, supra) # append gene_ids
summary(supra_df)
head(supra_df)
supraR <- dplyr::select(supra_df, gene_id, log2FoldChange, padj) # select relevant columns
head(supraR)
dim(supraR) #4475
supraR.vector <- dplyr::filter(supraR, padj < 0.05) # all the differentially expressed supra RL genes 
head(supraR.vector)
supraR.vector <- dplyr::select(supraR.vector, gene_id, log2FoldChange) #remove padj column
dim(supraR.vector)#539
head(supraR.vector)
names(supraR.vector)[2] <- "supraR_FC"
head(supraR.vector)

# start with the list of all relevant genes (the full list of deduped and trimmed list of genes)
R.vector <- left_join(gene_id, subR.vector, by = "gene_id")
dim(R.vector) #4475
head(R.vector)
tail(R.vector)
R.vector <- left_join(R.vector, supraR.vector, by = "gene_id")
head(R.vector)
dim(R.vector) #4475
head(R.vector, 10)
tail(R.vector, 10)

# build Venn Diagram input
venn_df <- R.vector %>%
  gather(key = exp_type, value = foldchange, -c(gene_id))
glimpse(venn_df)

## this is for positive fold change (upregulated for 22 relative to 27)
pos_venn_sub <- venn_df %>% 
  filter(exp_type == "subR_FC" & !is.na(foldchange) & foldchange > 0) %>% 
  pull(gene_id)

## this is for positive fold change (upregulated for 32 relative to 27)
pos_venn_supra <- venn_df %>% 
  filter(exp_type == "supraR_FC" & !is.na(foldchange) & foldchange < 0) %>% 
  pull(gene_id)

## this is for negative fold change (downregulated 22 relative to 27)
neg_venn_sub <- venn_df %>% 
  filter(exp_type == "subR_FC" & !is.na(foldchange) & foldchange < 0) %>% 
  pull(gene_id)

## this is for negative fold change (downregulated for 32 relative to 27)
neg_venn_supra <- venn_df %>% 
  filter(exp_type == "supraR_FC" & !is.na(foldchange) & foldchange > 0) %>% 
  pull(gene_id)
summary(pos_venn_sub)
summary(pos_venn_supra)
summary(neg_venn_sub)
summary(neg_venn_supra)
venn_list_pos <- list("22C" = pos_venn_sub,
                      "32C" = pos_venn_supra)
venn_list_neg <- list("22C" = neg_venn_sub,
                      "32C" = neg_venn_supra)
venn <- venn.diagram(venn_list_pos, NULL,
                          col = "transparent",
                          fill = c("#350498", "#FBA139"),
                          alpha = 0.50,
                          cex = 0.8,
                          fontfamily = "sans",
                          fontface = "bold",
                          cat.col = c("#350498", "#FBA139"),
                          cat.cex = 0.8,
                          cat.fontfamily = "sans",
                          margin = 0.2,
                          main.fontfamily = "sans",
                          print.mode = c("raw","percent"), # Show both numbers and percent
                          main.pos = c(0.5,0.9))
grid.draw(venn)
ggsave("vd_vOpt/VD_R_up_v27.pdf", plot = venn, height=120, width=120, units='mm', dpi=500)
dev.off()

## extract the overlap analysis so you have the list of genes that are in each section of the venn diagram
list <- calculate.overlap(venn_list_pos)
summary(list)

# core genes
write.csv(list$a3,"vd_vOpt/VD_R_upregulated_core.csv", row.names = TRUE)

core <- read_csv("vd_vOpt/VD_R_upregulated_core.csv")
summary(core) # should have 158; the "core" genes are included in this count
head(core)

# 22 vs 27 genes
write.csv(list$a1,"vd_vOpt/VD_R_upregulated_22v27.csv", row.names = TRUE)

sub <- read_csv("vd_vOpt/VD_R_upregulated_22v27.csv")
head(sub)
summary(sub)
head(core)
sub_df <- left_join(sub, core, by = "x")
sub_df
summary(sub_df)
head(sub_df)
sub_df <- sub_df[is.na(sub_df$...1.y),]
head(sub_df)
summary(sub_df)

write.csv(sub_df$x,"vd_vOpt/VD_R_upregulated_22v27_deduped.csv", row.names = TRUE)

# 32 vs 27 genes
write.csv(list$a2,"vd_vOpt/VD_R_upregulated_32v27.csv", row.names = TRUE)

supra <- read_csv("vd_vOpt/VD_R_upregulated_32v27.csv")
head(supra)
summary(supra)
head(core)
supra_df <- left_join(supra, core, by = "x")
supra_df
summary(supra_df)
head(supra_df)
supra_df <- supra_df[is.na(supra_df$...1.y),]
head(supra_df)
summary(supra_df)

write.csv(supra_df$x,"vd_vOpt/VD_R_upregulated_32v27_deduped.csv", row.names = TRUE)

# Downregulated R DEG Venn Diagrams - build Venn Diagram and extract gene lists
venn <- venn.diagram(venn_list_neg, NULL,
                          col = "transparent",
                          fill = c("#350498", "#FBA139"),
                          alpha = 0.50,
                          cex = 0.8,
                          fontfamily = "sans",
                          fontface = "bold",
                          cat.col = c("#350498", "#FBA139"),
                          cat.cex = 0.8,
                          cat.fontfamily = "sans",
                          margin = 0.2,
                          main.fontfamily = "sans",
                          print.mode = c("raw","percent"), # Show both numbers and percent
                          main.pos = c(0.5,0.9))
grid.draw(venn)
ggsave("vd_vOpt/VD_R_down_v27.pdf", plot = venn, height=120, width=120, units='mm', dpi=500)
dev.off()

## extract the overlap analysis so you have the list of genes that are in each section of the venn diagram
list <- calculate.overlap(venn_list_neg)
summary(list)

# core genes shared across day treatments
write.csv(list$a3,"vd_vOpt/VD_R_downregulated_core.csv", row.names = TRUE)

core <- read_csv("vd_vOpt/VD_R_downregulated_core.csv")
summary(core) # should have 436; the "core" genes are included in this count
head(core)

# 22 vs 27 genes
write.csv(list$a1,"vd_vOpt/VD_R_downregulated_22v27.csv", row.names = TRUE)

sub <- read_csv("vd_vOpt/VD_R_downregulated_22v27.csv")
head(sub)
summary(sub)
head(core)
sub_df <- left_join(sub, core, by = "x")
sub_df
summary(sub_df)
head(sub_df)
sub_df <- sub_df[is.na(sub_df$...1.y),]
head(sub_df)
summary(sub_df)

write.csv(sub_df$x,"vd_vOpt/VD_R_downregulated_22v27_deduped.csv", row.names = TRUE)

# 32 vs 27 genes
write.csv(list$a2,"vd_vOpt/VD_R_downregulated_32v27.csv", row.names = TRUE)

supra <- read_csv("vd_vOpt/VD_R_downregulated_32v27.csv")
head(supra)
summary(supra)
head(core)
supra_df <- left_join(supra, core, by = "x")
supra_df
summary(supra_df)
head(supra_df)
supra_df <- supra_df[is.na(supra_df$...1.y),]
head(supra_df)
summary(supra_df)

write.csv(supra_df$x,"vd_vOpt/VD_R_downregulated_32v27_deduped.csv", row.names = TRUE)
```


#cluster profiler replete 
```{r R_clusterProfiler, eval = TRUE, message = FALSE, warning = FALSE}

# Core
df <- read_csv("vd_vOpt/VD_R_upregulated_core.csv")
head(df)
df_ids <- df$x

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/22v32/R_upregulated_core_BP.csv", row.names = TRUE) # enricher results

# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/22v32/R_upregulated_core_MF.csv", row.names = TRUE) # enricher results

# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_CC, TERM2NAME = go2name_CC, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/22v32/R_upregulated_core_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/22v32/R_upregulated_core_KO.csv", row.names = TRUE) # enricher results

# 22 unique 
df <- read_csv("vd_vOpt/VD_R_upregulated_22v27.csv")
head(df)
df_ids <- df$x

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/22v32/R_upregulated_22_BP.csv", row.names = TRUE) # enricher results

# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/22v32/R_upregulated_22_MF.csv", row.names = TRUE) # enricher results

# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_CC, TERM2NAME = go2name_CC, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/22v32/R_upregulated_22_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/22v32/R_upregulated_22_KO.csv", row.names = TRUE) # enricher results

# 32 unique
df <- read_csv("vd_vOpt/VD_R_upregulated_32v27.csv")
head(df)
df_ids <- df$x

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/22v32/R_upregulated_32_BP.csv", row.names = TRUE) # enricher results

# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/22v32/R_upregulated_32_MF.csv", row.names = TRUE) # enricher results

# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_CC, TERM2NAME = go2name_CC, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/22v32/R_upregulated_32_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/22v32/R_upregulated_32_KO.csv", row.names = TRUE) # enricher results

# 22R v 27R down

# Core
df <- read_csv("vd_vOpt/VD_R_downregulated_core.csv")
head(df)
df_ids <- df$x

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/22v32/R_downregulated_core_BP.csv", row.names = TRUE) # enricher results

# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/22v32/R_downregulated_core_MF.csv", row.names = TRUE) # enricher results

# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_CC, TERM2NAME = go2name_CC, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/22v32/R_downregulated_core_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/22v32/R_downregulated_core_KO.csv", row.names = TRUE) # enricher results

# 22 unique 
df <- read_csv("vd_vOpt/VD_R_downregulated_22v27.csv")
head(df)
df_ids <- df$x

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/22v32/R_downregulated_22_BP.csv", row.names = TRUE) # enricher results

# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/22v32/R_downregulated_22_MF.csv", row.names = TRUE) # enricher results

# CC

enrich <- enricher(df_ids, TERM2GENE = go2gene_CC, TERM2NAME = go2name_CC, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/22v32/R_downregulated_22_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/22v32/R_downregulated_22_KO.csv", row.names = TRUE) # enricher results

# 32 unique
df <- read_csv("vd_vOpt/VD_R_downregulated_32v27.csv")
head(df)
df_ids <- df$x

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/22v32/R_downregulated_32_BP.csv", row.names = TRUE) # enricher results

# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/22v32/R_downregulated_32_MF.csv", row.names = TRUE) # enricher results

# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_CC, TERM2NAME = go2name_CC, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/22v32/R_downregulated_32_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/22v32/R_downregulated_32_KO.csv", row.names = TRUE) # enricher results
```

# venn diagram limited 
```{r L_venn, eval = TRUE, message = FALSE, warning = FALSE}

# set up limited dataframes
sub <- read_csv("../deseq2/outputs/L22vL27_DESeqresults.csv")
head(sub)

supra <- read_csv("../deseq2/outputs/L27vL32_DESeqresults.csv")
head(supra)

# sub 
sub_df <- cbind(gene_id, sub) # append gene_ids
summary(sub_df)
head(sub_df)
subL <- dplyr::select(sub_df, gene_id, log2FoldChange, padj) # select relevant columns
head(subL)
dim(subL) #4475
subL.vector <- dplyr::filter(subL, padj < 0.05) 
head(subL.vector)
subL.vector <- dplyr::select(subL.vector, gene_id, log2FoldChange) #remove padj column
dim(subL.vector) #1224
head(subL.vector)
names(subL.vector)[2] <- "subL_FC"
head(subL.vector)

# supra
supra_df <- cbind(gene_id, supra) # append gene_ids
summary(supra_df)
head(supra_df)
supraL <- dplyr::select(supra_df, gene_id, log2FoldChange, padj) # select relevant columns
head(supraL)
dim(supraL) #4475
supraL.vector <- dplyr::filter(supraL, padj < 0.05) # all the differentially expressed supra LL genes 
head(supraL.vector)
supraL.vector <- dplyr::select(supraL.vector, gene_id, log2FoldChange) #remove padj column
dim(supraL.vector) #1265
head(supraL.vector)
names(supraL.vector)[2] <- "supraL_FC"
head(supraL.vector)

# start with the list of all relevant genes (the full list of deduped and trimmed list of genes)
L.vector <- left_join(gene_id, subL.vector, by = "gene_id")
dim(L.vector) #4475
head(L.vector)
L.vector <- left_join(L.vector, supraL.vector, by = "gene_id")
head(L.vector)
dim(L.vector) #4475
head(L.vector, 10)
tail(L.vector, 10)

# build Venn Diagram input
venn_df <- L.vector %>%
  gather(key = exp_type, value = foldchange, -c(gene_id))
glimpse(venn_df)
# why is the fold change just a number that keeps going up?

## this is for positive fold change (upregulate for 22 relative to 27)
pos_venn_sub <- venn_df %>% 
  filter(exp_type == "subL_FC" & !is.na(foldchange) & foldchange > 0) %>% 
  pull(gene_id)

## this is for positive fold change (upregulate for 32 relative to 27)
pos_venn_supra <- venn_df %>% 
  filter(exp_type == "supraL_FC" & !is.na(foldchange) & foldchange < 0) %>% 
  pull(gene_id)

## this is for negative fold change (downregulate 22 relative to 27)
neg_venn_sub <- venn_df %>% 
  filter(exp_type == "subL_FC" & !is.na(foldchange) & foldchange < 0) %>% 
  pull(gene_id)

## this is for negative fold change (downregulate for 32 relative to 27)
neg_venn_supra <- venn_df %>% 
  filter(exp_type == "supraL_FC" & !is.na(foldchange) & foldchange > 0) %>% 
  pull(gene_id)
summary(pos_venn_sub) #672
summary(pos_venn_supra) # 652
summary(neg_venn_sub) #552
summary(neg_venn_supra) #613

venn_list_pos <- list("22C" = pos_venn_sub,
                      "32C" = pos_venn_supra)
venn_list_neg <- list("22C" = neg_venn_sub,
                      "32C" = neg_venn_supra)
venn <- venn.diagram(venn_list_pos, NULL,
                          col = "transparent",
                          fill = c("#350498", "#FBA139"),
                          alpha = 0.50,
                          cex = 0.8,
                          fontfamily = "sans",
                          fontface = "bold",
                          cat.col = c("#350498", "#FBA139"),
                          cat.cex = 0.8,
                          cat.fontfamily = "sans",
                          margin = 0.2,
                          main.fontfamily = "sans",
                          print.mode = c("raw","percent"), # Show both numbers and percent
                          main.pos = c(0.5,0.9))

grid.draw(venn)
ggsave("vd_vOpt/VD_L_up_v27.pdf", plot = venn, height=120, width=120, units='mm', dpi=500)
dev.off()

## extract the overlap analysis so you have the list of genes that are in each section of the venn diagram
list <- calculate.overlap(venn_list_pos)
summary(list)

# core genes
# this is saying there are no core genes, which is weird because there are core genes in the venn diagram
write.csv(list$a3,"vd_vOpt/VD_L_upregulated_core.csv", row.names = TRUE)

core <- read_csv("vd_vOpt/VD_L_upregulated_core.csv")
summary(core) # should have 319; the "core" genes are included in this count
head(core)

# 22 vs 27 genes
write.csv(list$a1,"vd_vOpt/VD_L_upregulated_22v27.csv", row.names = TRUE)

sub <- read_csv("vd_vOpt/VD_L_upregulated_22v27.csv")
head(sub)
summary(sub)
head(core)
sub_df <- left_join(sub, core, by = "x")
sub_df
summary(sub_df)
head(sub_df)
sub_df <- sub_df[is.na(sub_df$...1.y),]
# the above line is finding the difference between the core and the core + sub
head(sub_df)
summary(sub_df)
dim(sub_df) #353
write.csv(sub_df$x,"vd_vOpt/VD_L_upregulated_22v27_deduped.csv", row.names = TRUE)

# 32 vs 27 genes
write.csv(list$a2,"vd_vOpt/VD_L_upregulated_32v27.csv", row.names = TRUE)

supra <- read_csv("vd_vOpt/VD_L_upregulated_32v27.csv")
head(supra)
summary(supra)
head(core)
supra_df <- left_join(supra, core, by = "x")
supra_df
summary(supra_df)
head(supra_df)
supra_df <- supra_df[is.na(supra_df$...1.y),]
dim(supra_df) #333
head(supra_df)
summary(supra_df)

write.csv(supra_df$x,"vd_vOpt/VD_L_upregulated_32v27_deduped.csv", row.names = TRUE)

# Downregulated Day L DEG Venn Diagrams - build Venn Diagram and extract gene lists
venn <- venn.diagram(venn_list_neg, NULL,
                          col = "transparent",
                          fill = c("#350498", "#FBA139"),
                          alpha = 0.50,
                          cex = 0.8,
                          fontfamily = "sans",
                          fontface = "bold",
                          cat.col = c("#350498", "#FBA139"),
                          cat.cex = 0.8,
                          cat.fontfamily = "sans",
                          margin = 0.2,
                          main.fontfamily = "sans",
                          print.mode = c("raw","percent"), # Show both numbers and percent
                          main.pos = c(0.5,0.9))
grid.draw(venn)
ggsave("vd_vOpt/VD_L_down_v27.pdf", plot = venn, height=120, width=120, units='mm', dpi=500)
dev.off()

## extract the overlap analysis so you have the list of genes that are in each section of the venn diagram
list <- calculate.overlap(venn_list_neg)
summary(list)

# core genes shared across day treatments
write.csv(list$a3,"vd_vOpt/VD_L_downregulated_core.csv", row.names = TRUE)

core <- read_csv("vd_vOpt/VD_L_downregulated_core.csv")
summary(core) # should have 79; the "core" genes are included in this count
head(core)

# 21 vs 27 genes
write.csv(list$a1,"vd_vOpt/VD_L_downregulated_22v27.csv", row.names = TRUE)

sub <- read_csv("vd_vOpt/VD_L_downregulated_22v27.csv")
head(sub)
summary(sub)
head(core)
sub_df <- left_join(sub, core, by = "x")
sub_df
summary(sub_df)
head(sub_df)
sub_df <- sub_df[is.na(sub_df$...1.y),]
dim(sub_df) #280
head(sub_df)
summary(sub_df)

write.csv(sub_df$x,"vd_vOpt/VD_L_downregulated_22v27_deduped.csv", row.names = TRUE)

# 32 vs 27 genes
write.csv(list$a2,"vd_vOpt/VD_L_downregulated_32v27.csv", row.names = TRUE)

supra <- read_csv("vd_vOpt/VD_L_downregulated_32v27.csv")
head(supra)
summary(supra)
head(core)
supra_df <- left_join(supra, core, by = "x")
supra_df
summary(supra_df)
head(supra_df)
supra_df <- supra_df[is.na(supra_df$...1.y),]
dim(supra_df) #341
head(supra_df, 10)
tail(supra_df, 10)
summary(supra_df)

write.csv(supra_df$x,"vd_vOpt/VD_L_downregulated_32v27_deduped.csv", row.names = TRUE)
```


# limited cluster profiler 
```{r L_clusterProfiler, eval = TRUE, message = FALSE, warning = FALSE}

# Core
df <- read_csv("vd_vOpt/VD_L_upregulated_core.csv")
head(df)
df_ids <- df$x

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/vOpt/L_upregulated_core_BP.csv", row.names = TRUE) # enricher results

# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/vOpt/L_upregulated_core_MF.csv", row.names = TRUE) # enricher results

# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_CC, TERM2NAME = go2name_CC, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/vOpt/L_upregulated_core_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/vOpt/L_upregulated_core_KO.csv", row.names = TRUE) # enricher results

# 22 unique 
df <- read_csv("vd_vOpt/VD_L_upregulated_22v27.csv")
head(df)
df_ids <- df$x

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/vOpt/L_upregulated_22_BP.csv", row.names = TRUE) # enricher results

# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/vOpt/L_upregulated_22_MF.csv", row.names = TRUE) # enricher results

# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_CC, TERM2NAME = go2name_CC, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/vOpt/L_upregulated_22_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/vOpt/L_upregulated_22_KO.csv", row.names = TRUE) # enricher results

# 32 unique
df <- read_csv("vd_vOpt/VD_L_upregulated_32v27.csv")
head(df)
df_ids <- df$x

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/vOpt/L_upregulated_32_BP.csv", row.names = TRUE) # enricher results

# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/vOpt/L_upregulated_32_MF.csv", row.names = TRUE) # enricher results

# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_CC, TERM2NAME = go2name_CC, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/vOpt/L_upregulated_32_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/vOpt/L_upregulated_32_KO.csv", row.names = TRUE) # enricher results

# 22LD v 27LD down
# Core
df <- read_csv("vd_vOpt/VD_L_downregulated_core.csv")
head(df)
df_ids <- df$x

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/vOpt/L_downregulated_core_BP.csv", row.names = TRUE) # enricher results

# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/vOpt/L_downregulated_core_MF.csv", row.names = TRUE) # enricher results

# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_CC, TERM2NAME = go2name_CC, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/vOpt/L_downregulated_core_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/vOpt/L_downregulated_core_KO.csv", row.names = TRUE) # enricher results

# 22 unique 
df <- read_csv("vd_vOpt/VD_L_downregulated_22v27.csv")
head(df)
df_ids <- df$x

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/vOpt/L_downregulated_22_BP.csv", row.names = TRUE) # enricher results

# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/vOpt/L_downregulated_22_MF.csv", row.names = TRUE) # enricher results

# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_CC, TERM2NAME = go2name_CC, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/vOpt/L_downregulated_22_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/vOpt/L_downregulated_22_KO.csv", row.names = TRUE) # enricher results

# 32 unique
df <- read_csv("vd_vOpt/VD_L_downregulated_32v27.csv")
head(df)
df_ids <- df$x

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/vOpt/L_downregulated_32_BP.csv", row.names = TRUE) # enricher results

# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/vOpt/L_downregulated_32_MF.csv", row.names = TRUE) # enricher results

# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_CC, TERM2NAME = go2name_CC, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/vOpt/L_downregulated_32_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/vOpt/L_downregulated_32_KO.csv", row.names = TRUE) # enricher results
```



### Temperatures (22 v 32) R and L cluster profiler
```{r 22RDv32RD_clusterProfiler, eval = TRUE, message = FALSE, warning = FALSE}

#22R vs 32R
# up
df <- read_csv("../cluster_profiler/vectors/R22vR32_up_vector.csv")
head(df)
df_ids <- df$gene_id

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/22v32/R22vR32_up_BP.csv", row.names = TRUE) # enricher results

# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/22v32/R22vR32_up_MF.csv", row.names = TRUE) # enricher results

# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_CC, TERM2NAME = go2name_CC, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/22v32/R22vR32_up_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/22v32/R22vR32_up_KO.csv", row.names = TRUE) # enricher results


# down
df <- read_csv("../cluster_profiler/vectors/R22vR32_down_vector.csv")
head(df)
df_ids <- df$gene_id

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/22v32/R22vR32_down_BP.csv", row.names = TRUE) # enricher results

# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/22v32/R22vR32_down_MF.csv", row.names = TRUE) # enricher results

# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_CC, TERM2NAME = go2name_CC, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/22v32/R22vR32_down_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/22v32/R22vR32_down_KO.csv", row.names = TRUE) # enricher results


#22L vs 32L
# up
df <- read_csv("../cluster_profiler/vectors/L22vL32_up_vector.csv")
head(df)
df_ids <- df$gene_id

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/22v32/L22vL32_up_BP.csv", row.names = TRUE) # enricher results

# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/22v32/L22vL32_up_MF.csv", row.names = TRUE) # enricher results

# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_CC, TERM2NAME = go2name_CC, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/22v32/L22vL32_up_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/22v32/L22vL32_up_KO.csv", row.names = TRUE) # enricher results

# down
df <- read_csv("../cluster_profiler/vectors/L22vL32_down_vector.csv")
head(df)
df_ids <- df$gene_id

# BP
enrich <- enricher(df_ids, TERM2GENE = go2gene_BP, TERM2NAME = go2name_BP, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/22v32/L22vL32_down_BP.csv", row.names = TRUE) # enricher results

# MF
enrich <- enricher(df_ids, TERM2GENE = go2gene_MF, TERM2NAME = go2name_MF, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/22v32/L22vL32_down_MF.csv", row.names = TRUE) # enricher results

# CC
enrich <- enricher(df_ids, TERM2GENE = go2gene_CC, TERM2NAME = go2name_CC, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/22v32/L22vL32_down_CC.csv", row.names = TRUE) # enricher results

# KO
enrich <- enrichKEGG(gene = df_ids, organism = "ko", universe = bg_KO, minGSSize = 5)
enrich

write.csv(enrich, file="../cluster_profiler/22v32/L22vL32_down_KO.csv", row.names = TRUE) # enricher results
```

